---
title: "Assignment 1"
author: "Rogelio Cadena"
date: "9/12/2020"
output:
 html_document:
  theme: lumen
  toc: true
  toc_depth: 3
  toc_float: true
---

```{r load libraries}
library(tidycensus)
library(tidyverse)
library(viridis)
library(ggplot2)
library(ggthemes)
```

```{r}
vars <- c("B06009_004",
          "B25001_001",
          "B25003_003",
          "B06007_002",
          "B01002_001",
          "B01003_001",
          "B03002_003",
          "B03002_004",
          "B03002_006",
          "B03002_012",
          "B01001_026",
          "B06010_003",
          "B06010_004",
          "B06010_005",
          "B06010_006",
          "B06010_007",
          "B06010_008",
          "B06010_009",
          "B06010_010",
          "B06010_011")
```


```{r, message=FALSE}
IL_tracts <- get_acs(geography = "tract", variables = vars,
                           state = "IL", county = "Cook", output = "wide") %>% 
             rename(tot_bach = "B06009_004E",
                    tot_hh = "B25001_001E",
                    tot_rent_occ = "B25003_003E",
                    tot_eng = "B06007_002E",
                    med_age = "B01002_001E",
                    tot_pop = "B01003_001E",
                    pop_inc = "B06010_003E",
                    bel_10 = "B06010_004E", 
                    bel_15 = "B06010_005E",
                    bel_25 = "B06010_006E",
                    bel_35 = "B06010_007E",
                    bel_50 = "B06010_008E",
                    bel_65 = "B06010_009E",
                    bel_75 = "B06010_010E",
                    abo_75 = "B06010_011E",
                    nh_white_pop = "B03002_003E",
                    nh_black_pop = "B03002_004E",
                    nh_asian_pop = "B03002_006E",
                    hispanic_pop = "B03002_012E") %>%
              mutate(pct_rent_occ = tot_rent_occ / tot_hh,
                     pct_bach = tot_bach / tot_pop,
                     pct_eng = tot_eng / tot_pop,
                     majority_inc = case_when(
                     (bel_10 + bel_15 + bel_25) / pop_inc > 0.5 ~ "bel_25",
                     (bel_35 + bel_50 + bel_65 + abo_75) / pop_inc > 0.5 ~ "abo_25",
                     (bel_10 + bel_15 + bel_25 + bel_35 + bel_50 + bel_65 +
                      bel_75 + abo_75)/pop_inc < 0.5 ~ "other", 
                     TRUE ~ "None"),
                     majority_race_eth = case_when(
                     nh_white_pop / tot_pop > 0.5 ~ "nh_white",
                     nh_black_pop / tot_pop > 0.5 ~ "nh_black",
                     nh_asian_pop / tot_pop > 0.5 ~ "nh_asian",
                    hispanic_pop / tot_pop > 0.5 ~ "hispanic",
                    (nh_white_pop + nh_black_pop + nh_asian_pop + hispanic_pop) /
                     tot_pop < 0.5 ~ "other",
                     TRUE ~ "None")) %>%
              filter(tot_pop > 0, tot_hh > 0) %>%
              select(GEOID, pct_rent_occ, pct_bach, pct_eng,
                     majority_inc, med_age, majority_race_eth, tot_pop)
```

#### Getting data from the 2014-2018 5-year ACS

# Plot 1
#### Plotting a majority of the variables

```{r}
ggplot(IL_tracts, 
       aes(x = tot_pop,
           y = med_age,
           shape = majority_inc,
           color = majority_race_eth,
           alpha = pct_eng)) +
  geom_point() +
  stat_smooth(method = "lm", aes(fill=majority_race_eth)) +
  scale_size_continuous(name = "Tract population") +
  scale_x_continuous(name = "Total Population",
                     trans = "log",
                     breaks =c(1000,2500,5000,15000),
                     labels = c(1000,2500,5000,15000)) +
  scale_y_continuous(name = "Median age within tract",
                     breaks = seq(10, 80, by = 10)) +
  scale_color_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-Hispanic Asian",
                                  "Non-Hispanic White",
                                  "Non-Hispanic Black", 
                                  "No Majority")) +
  scale_shape(name = element_blank(), 
              labels = c("Income Below $25,000", "Income Above $25,000")) +
  theme_test() +
  theme(legend.key.size = unit(0.1, "cm"))
```


# Plot 2
#### Scatter plot of percent of individuals that only speak English and percent with bachelor’s degree

```{r}
ggplot(IL_tracts,
  aes(pct_eng,pct_bach)
        )+
  geom_point() + 
  geom_smooth(method = "lm") +
  geom_jitter() + 
  labs(
    x='Percent of individuals that only speak English',
    y='Percent of individuals with a Bachelors Degree',
    title="",
    caption = ""
  ) +
  theme_igray()
```

# Plot 3
#### Scatter plot of percent with bachelor’s degree and percent of individuals that only speak English by Race/Ethnicity

```{r}
ggplot(IL_tracts, aes(x=pct_eng, y=pct_bach, size = tot_pop, fill = majority_race_eth)) +
    geom_point(alpha=0.5, shape=21, color="black") + 
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="B") +
    scale_x_continuous(name = "Percent of individuals that only speak English",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
    scale_y_continuous(name = "Percent with bachelor's degree",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  coord_flip() +
    theme_calc() 
```


# Plot 4
#### Bar chart of majority income and percent of individuals with a Bachelor's Degree

```{r}
ggplot(IL_tracts, aes(x=majority_inc, y=pct_bach, fill = majority_race_eth)) +
  geom_bar(position = "dodge", stat = "identity", width = 0.5) + 
  scale_y_continuous(name = "Percent of individuals with a Bachelor's Degree",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_x_discrete(name = "Majority income",
                       labels = c("Below $25,000",
                                  "Above $25,000",
                                  "No data")) +
 
  theme_bw()
```

# Plot 5
#### Bar chart of majority race/ethnicity and percent of renter occupied housing

```{r}
ggplot(IL_tracts, aes(x=majority_race_eth,
                      y=pct_rent_occ,
                      color=majority_race_eth)) +
  geom_point(position = "jitter",
             size = 0.35,
             alpha = 0.35) + 
  geom_segment( aes(x=majority_race_eth, xend=majority_race_eth, y=0, yend=pct_rent_occ)) + 
  scale_y_continuous(name = "Percent of renter occupied housing",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Hispanic",
                                  "Non-H Asian",
                                  "Non-H White",
                                  "Non-H Black", 
                                  "No Majority")) +
  coord_flip() +
  theme_grey()
```

# Plot 6
#### Box plot of percent with a bachelor’s degree, and majority race

```{r}
 ggplot(IL_tracts, aes(x=majority_race_eth, y=pct_bach, fill = majority_race_eth)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="B") +
    geom_jitter(color="blue", size=0.4, alpha=0.9) +
    scale_y_continuous(name = "Percent with bachelor's degree",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
    scale_x_discrete(name = "Majority Race",
                       labels = c("Hispanic",
                                  "Non-H Asian",
                                  "Non-H White",
                                  "Non-H Black", 
                                  "No Majority")) +
  theme_stata()
```

# Plot 7
#### Violin plot of percent only English, and majority income

```{r}
ggplot(IL_tracts, aes(x = majority_inc, y = pct_eng, fill = majority_inc)) +
    geom_violin(width=2.1, size=0.2, alpha = 0.65) +
    scale_color_viridis(discrete=TRUE) +
    scale_y_continuous(name = "Percent of individuals that only speak English",
                       breaks = seq(0, 1, by = 0.25),
                       labels = paste(seq(0, 100, by = 25), "%", sep = "")) +
  scale_x_discrete(name = "Majority Income",
                       labels = c("Above $25k",
                                  "Below $25K",
                                  "None")) +
  scale_fill_discrete(name = "Majority Income",
                               labels = c("Below $25K",
                                  "Above $25K",
                                  "No data")) +
    theme_light()
```

# Plot 8
#### Pie Chart of Majority Income

```{r}
pie_chart <- ggplot (IL_tracts, aes (x=1, fill = majority_inc)) +
  geom_bar(position = "fill") +
  coord_polar(theta = "y") +
  scale_y_continuous(name = "Number of tracts") +
  scale_fill_discrete(name = "Majority Income",
                               labels = c("Below $25,000",
                                  "Above $25,000",
                                  "No data")) +
                                         
  theme_void()

pie_chart
```

# Plot 9
#### Pie Chart of Bachelor's Degree attainment by Race/Ethnicity

```{r}
ggplot(IL_tracts,
  aes(majority_race_eth,pct_bach)
        )+
  geom_bar(stat="identity", fill= "orange")+   
  
  labs(
    x= NULL,
    y= 'Median Age'
  ) +
  coord_polar("x", start = 0) +
  theme_hc()
```

# Plot 10
#### Area chart of percent of rental occupation

```{r}
ggplot(IL_tracts, aes(x=pct_rent_occ)) +
  geom_density(fill="brown", alpha=0.45) +
  scale_x_continuous(name = "Percent of rental occupation",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  theme_classic()
```
   