---
title: "Assignment 2"
author: "Rogelio Cadena"
date: "9/20/2020"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r load libraries, results = "hide", message = FALSE, warning = FALSE}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
```

## Mapping India
#### I am plotting India because I have great appreciation for the people and culture
```{r message=FALSE}
india_states <- ne_states(country = "India", returnclass = "sf")
india_map <- ggplot(india_states) +
  geom_sf() +
  theme_bw()
india_map +
  annotation_scale()
```

## Projections
```{r}
India_Albers <- "+proj=aea +lat_1=28 +lat_2=12 +lat_0=20 +lon_0=78 +x_0=2000000 +y_0=2000000 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
India_EastNAD83 <- "+proj=tmerc +lat_0=37.5 +lon_0=-85.66666666666667 +k=0.999966667 +x_0=99999.99989839978 +y_0=249364.9987299975 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"
India_GoogleM <- "+proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +wktext  +no_defs"
```

## The India Albers Equal Area Projection
```{r}
ggplot(india_states) +
  geom_sf() +
  coord_sf(crs = India_Albers) +
  theme_wsj()
```

## The NAD83 India East (ftUS) Projection
```{r}
ggplot(india_states) +
  geom_sf() +
  coord_sf(crs = India_EastNAD83) +
  theme_wsj()
```

## The Google Mercator India Projection
```{r}
ggplot(india_states) +
  geom_sf() +
  coord_sf(crs = India_GoogleM) +
  theme_wsj()
```

## Transforming data in a projected coordinate system
```{r}
states_transformed <- india_states %>%
  st_transform(crs = India_Albers)
```

## Representing Covid-19 cases in India by state
#### The dataset for Covid cases in India was acquired through data.world and originally is sourced from the Ministry of Health & Family Welfare in India
```{r}
covid_ind <- read.csv("india_covid_19_state_wise_info1.csv")

covid_states<- states_transformed %>% 
  left_join(covid_ind, by = c("name" = "state_ut")) %>%
  select(confirmed_cases, active_cases, recovered_cases, death_cases)
```

## Cloropeth Map
#### This map highlights states in the country with the highest to lowest confirmed cases of Covid cases from a range of 70,0000 - 10,000 cases
```{r}
ggplot(covid_states, aes(fill = confirmed_cases)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "Number of\nconfirmed\nCOVID-19\ncases as of\nMay 29, 2020",
    breaks = seq(10000, 70000, by = 10000),
    labels = formatC(seq(10000, 70000, by = 10000), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_void() +
  theme(legend.background = element_blank())
```

### Continuous Cartogram
#### This map distorts state sizes according to their relative number of Covid- 19 cases. For example, the state Maharashstra (colored in yellow), is represented much large than its actual size due to the number of confirmed cases being the highest in the country.
```{r, message=FALSE, warning=FALSE}
covid_cartogram_ncont <- covid_states %>%
  cartogram_cont("confirmed_cases")
ggplot(covid_cartogram_ncont, aes(fill = confirmed_cases)) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(
    name = "Number of\nconfirmed\nCOVID-19 cases\nas of May 29, 2020",
    breaks = seq(10000, 70000, by = 10000),
    labels = formatC(seq(10000, 70000, by = 10000), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_void() +
  theme(legend.background = element_blank())
```

## Proportional Symbol Map
#### This same reading can also be depicted as circular symbols where the various sizes relate to the number of confirmed cases.
```{r, message = FALSE, warning=FALSE}
covid_centroids <- covid_states %>%
  st_centroid()
ggplot(states_transformed) +
  geom_sf(fill = NA, color = "gray") +
  geom_sf(data = covid_centroids, 
          aes(size = confirmed_cases),
          alpha = 0.65, color = "pink") +
  
  scale_size_continuous(
    name = "Number of confirmed\nCOVID-19 cases\nas of May 29, 2020",
    breaks = seq(10000, 70000, by = 10000),
    labels = formatC(seq(10000, 70000, by = 10000),  
                     big.mark = ",", format = "f", digits = 0),
    range = c(0, 10)) +
  theme_tufte() +
  theme(legend.background = element_blank())
```

