---
title: "Assignment 3"
author: "Rogelio Cadena"
date: "9/27/2020"
output: 
  html_document:
  theme: cosmo
toc: true
toc_depth: 3
toc_float: true
---

```{r message=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
library(ggplot2)
```

# Loading the Data
#### I have chosen to use four datasets (2 polygons and 2 points) showing neighborhoods, parks, grocery stores and cta bus stops
```{r}
nhoods <- st_read("https://data.cityofchicago.org/api/geospatial/bbvz-uum9?method=export&format=KML", 
                  quiet = TRUE)

parks <- st_read("https://data.cityofchicago.org/api/geospatial/ej32-qgdr?method=export&format=KML", 
                  quiet = TRUE)

mural <- st_read("https://data.cityofchicago.org/api/geospatial/we8h-apcf?method=export&format=KML", 
                  quiet = TRUE)

bus_stops <- st_read("CTA_BusStops.kml", 
                  quiet = TRUE)
```

# Transforming the data
#### I've transformed the data to the NAD 1983 East Illinois State Plane FIPS 1201 Feet 
```{r}
IL_state_plane <- "+proj=tmerc +lat_0=36.66666666666666 +lon_0=-88.33333333333333 +k=0.9999749999999999 +x_0=300000 +y_0=0 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"

nhoods <- nhoods %>%
  st_transform(IL_state_plane)

parks <- parks %>%
  st_transform(IL_state_plane)

mural <- mural %>%
  st_transform(IL_state_plane)

bus_stops <- bus_stops %>%
  st_transform(IL_state_plane)
```

# Creating a buffer
#### I've transformed the data to the NAD 1983 East Illinois State Plane System 
```{r}
ggplot(bus_stops) +
  geom_sf(color = "blue", size= 0.01) +
  geom_sf(data = mural, color = "orange", size = 0.05) +
  geom_sf(data = parks, color = "darkgreen") +
  geom_sf(data = nhoods, color= "gray", fill= NA) +
  theme_map() +
  annotation_scale()
```

# 1 Counting Points in a Polygon
#### The number and proportion of A points (Bus Stops) within C polygons (Neighborhoods)
```{r}
nhoods <- nhoods %>%
  mutate(num_bus_stops = lengths(st_covers(nhoods, bus_stops)))

ggplot(nhoods) +
  geom_sf(color = NA, 
          aes(fill = num_bus_stops)) +
  scale_fill_viridis_c(name = "Chicago neighborhoods\nby number of bus stops",
                       breaks = breaks <- seq(0, 350, by =50),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "bus stops")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

# 2 Counting Points in a Polygon
#### The average number of A points (Bus Stops) within C polygons (Neighborhoods)
```{r}
nhoods <- nhoods %>%
  mutate(area = set_units(st_area(nhoods), km^2)) %>%
  mutate(bus_stops_dens = as.numeric(num_bus_stops / area))

ggplot(nhoods) +
  geom_sf(color = NA,
          aes(fill = bus_stops_dens)) +
  scale_fill_viridis_c(name = 
                         "Chicago neighborhoods\nby Bus Stop Density",
                       breaks = breaks <- seq(1, 250, by = 25),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "bus stops per square km")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  labs(title = "Density of Bus Stops in Chicago Neighborhoods") +
  theme_map() +
    theme(legend.background = element_rect(fill=alpha("white", 0.5), color = "gray"))
```


# 3 Counting Points in a Polygon
#### The number and proportion of B points (Murals) within C polygons (Neighborhoods)
```{r message=FALSE}
nhoods <- nhoods %>%
  mutate(num_mural = lengths(st_covers(nhoods, mural)))

ggplot(nhoods) +
  geom_sf(color = NA, 
          aes(fill = num_mural)) +
  scale_fill_viridis_c(name = "Chicago neighborhoods\nby number of murals",
                       breaks = breaks <- seq(0, 50, by =10),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "mural")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

# 4 Identifying overlapping polygons
#### The number and proportion of C polygons (Neighborhoods) that overlap with D polygons (Parks)
```{r}
nhoods <- nhoods %>%
  mutate(num_parks = lengths(st_overlaps(nhoods, parks))) %>%
  mutate(has_parks = num_parks > 0)

n_parks_nhoods <- sum(nhoods$has_parks)

n_parks_nhoods
```
```{r message=FALSE}
left_side  <- st_bbox(parks)$xmin
top_side <- st_bbox(parks)$ymax

ggplot(parks) +
  geom_sf(fill = "darkgreen", color = NA) +
  geom_sf(data = nhoods,
          aes(fill = has_parks)) +
  scale_fill_manual(values = c("cornsilk1", "darkgreen"),
          name = "Chicago Neighborhoods\nby presence of a park", 
          labels = c("Neighborhood without\nan overlapping park",
                     "Neighborhood with an\noverlapping park")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side,
           y = top_side - 1000, 
           label = paste(n_parks_nhoods ,
                         "of Chicago's", 
                         length(nhoods$Name),
                         "neighborhoods contain\nor overlap with", 
                         "a park."),
           hjust = 0.4, vjust = 0.3, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "gray"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```


# 5 Counting Polygon in a Polygon
#### The number and proportion of D polygons (Parks) that overlap with  C polygons (Neighborhoods)
```{r message=FALSE}
nhoods <- nhoods %>%
  mutate(num_parks = lengths(st_covers(nhoods, parks)))

ggplot(nhoods) +
  geom_sf(color = NA, 
          aes(fill = num_parks)) +
  scale_fill_viridis_c(name = "Chicago neighborhoods\nby number of parks",
                       breaks = breaks <- seq(0, 20, by =5),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "parks")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

# 6 Subsetting points within a polygon
#### The number and proportion of B points (murals) with a specified distance of D polygons (parks)
```{r}
parks_buffer <- st_buffer(parks, dist = 400) %>%
  st_union()

ggplot(parks_buffer) +
  geom_sf() +
  theme_map()

bus_parks <- bus_stops[parks_buffer,]
  
ggplot(parks_buffer) +
  geom_sf() +
  geom_sf(data = bus_parks, 
          color = "blue", 
          size = 0.01) +
  theme_map()
```
```{r}
bus_stops <- bus_stops %>%
  st_join(bus_parks) %>%
  mutate(by_parks = !is.na(Name.y))
```

```{r}
n_parks_bus_stops <- sum(bus_stops$by_parks)

n_parks_bus_stops
```


```{r message=FALSE}
left_side <- st_bbox(bus_stops)$xmin
top_side <- st_bbox(bus_stops)$ymax

ggplot(parks) +
  geom_sf(fill = "darkgreen", color = NA) +
  geom_sf(data = bus_stops, size = 0.01,
          aes(color = by_parks)) +
  scale_color_manual(values = c("lightblue", "darkblue"),
          name = "Chicago bus stops\nby distance to parks", 
          labels = c("No bus stop within 400 m",
                     "Bus stops within 400 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_bus_stops, big.mark = ","),
                         "bus stops in Chicago\n", 
                         prettyNum(n_bus_stops, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_parks_bus_stops, digits = 0),
                         "%) are within 400\nmeters of parks.",
                         sep = ""),
           hjust = 0, vjust = .6, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```
